---
title: "Project_1_Reproducible_Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: taking out the NAs
This section loads the necessary libraries in R and also reads the data from the file.
```{r}
library(dplyr)
library(tidyr)
library(cowplot)

data_missing_values <- read.csv("activity.csv")
activity <- data_missing_values[complete.cases(data_missing_values$steps),]
```

## Part 2: mean number of steps per day
This section groups the data frame and plots the data, while also giving the mean and
median steps per day.
```{r}
days_count <- group_by(activity, date)
x <- summarize(days_count, total_steps = sum(steps))
ggplot(x, aes(total_steps)) + geom_histogram(binwidth = 1000)
ggsave("plot1.png", path = "./figure/", height = 5, width = 5)

mean_steps_day <- mean(x$total_steps)
median_steps_day <- median(x$total_steps)

mean_steps_day
median_steps_day
```

## Part 3: average daily activity pattern
This section groups the data frame and plots the data, while also giving the max interval mean.
```{r}
average_daily <- group_by(activity, interval)
y <- summarize(average_daily, interval_mean = mean(steps))
ggplot(y, aes(interval, interval_mean)) + geom_line()
ggsave("plot2.png", path = "./figure/", height = 5, width = 5)

max_interval <- y$interval[which.max(y$interval_mean)]
max_interval
```

## Part 4: inserting missing values
This section imputes values into the data where there is NA values and then does the
same analysis that is done in Part 2.
```{r}
is_NA <- is.na(data_missing_values$steps)
rows_with_NA <- sum(is_NA)

imputed <- data_missing_values
df_NA <- imputed[is.na(imputed$steps),]

map_NA <- which(is_NA)
for (i in 1:length(df_NA$interval)) {
  imputed[map_NA[i], 1] <- filter(y, interval == df_NA$interval[i])$interval_mean
}

days_count_imp <- group_by(imputed, date)
z <- summarize(days_count_imp, total_steps = sum(steps))

new_x <- mutate(x, "dataset" = "no-NA")
new_z <- mutate(z, "dataset" = "imputed")

new_df <- rbind(new_x, new_z)

ggplot(new_df, aes(total_steps, fill = dataset)) +
  geom_histogram(binwidth = 1000, position = "identity")
ggsave("plot3.png", path = "./figure/", height = 5, width = 5)

mean_steps_day_imp <- mean(z$total_steps)
median_steps_day_imp <- median(z$total_steps)

mean_steps_day_imp
median_steps_day_imp
```

## Part 5: differences between the weekday and the weekend
This section compares weekend activity to weekday activity.
```{r}
imputed2 <- mutate(imputed, "day_type" = as.factor("weekday"))
levels(imputed2$day_type) <- c("weekday", "weekend")
imputed2$date <- as.Date(imputed2$date)

imputed2 <- within(imputed2, day_type[weekdays(date) == c("Saturday", "Sunday")] <- 
                     as.factor("weekend"))

days_count_imp2 <- group_by(imputed2, interval, day_type)
compare <- summarize(days_count_imp2, interval_mean = mean(steps))

ggplot(compare, aes(interval, interval_mean)) + geom_line() +
  facet_wrap(~day_type, nrow = 2)
ggsave("plot4.png", path = "./figure/", height = 5, width = 7)
```


